Contextuality and the Born rule in fuzzy quantum theories

### 2017-03-26 prop_letter.tex
prop_letter.tex Updated by Prof. Gerardo Ortiz...

In conventional quantum theory, when we measure a property $\mathbf{O}$
of a physical system whose state is represented by a density matrix~$\rho$,
the probability to obtain the value $\lambda$ of that property is determined by the
Born rule $\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{\lambda}\right)=\Tr\left(\rho P_{\lambda}\right)$,
where $P_{\lambda}$ denotes the projector onto the eigenspace corresponding
to the eigenvalue $\lambda$ of the Hermitian operator
$\mathbf{O}$~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. 
Alternatively, one may say that the state $\rho$ determines the 
probability measure $\mu_{\rho}^{\mathrm{\mathrm{B}}}$ defined as a
function from the set $\events$ of all possible projectors to the interval 
$[0,1]$, and satisfies the following probability axioms~
\citep{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}: 
\begin{itemize}
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\mathbb{0})=0$. 
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\mathbb{1})=1$. 
\item For any projection $P$, $\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\mathbb{1}-P\right)=1-\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\sum_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{i}\right)\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique mixed state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{\mathrm{B}}}$
in a Hilbert space of dimension $d\geq3$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.

\paragraph{Quantum Interval-valued Probability Measure (IVPM)}

In order to model the idea of finite precision measurement, we replace
the infinity precise~$[0,1]$ by a collection of finite number of
intervals~$\mathscr{I}$, and consider quantum IVPMs~$\bar{\mu}:\events\rightarrow\mathscr{I}$.
Based on the intuitive requirement, $\left[0,0\right]$, $\left[1,1\right]\in\mathscr{I}$,
and operations on intervals,\begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & \left[1,1\right]-\left[l,r\right]=\left[1-r,1-l\right]\textrm{ and}\\
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]
\end{eqnarray}
\end{subequations}a quantum IVPM~$\bar{\mu}$ should satisfy corresponding
axioms:
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item For any projection $P$, 
\begin{equation}
\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Complement}
\end{equation}
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
Notice that Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
relaxes the equality in Eq.~(\ref{eq:QuantumProbability-Addition})
to inclusion because the equality is usually too strong to hold. For
example, if we consider $\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $
and $\bar{\mu}\left(P\right)=\left[l,r\right]$, by Eq.~(\ref{eq:QuantumInterval-valuedProbability-Complement})
and (\ref{eq:interval-operations}), we have 
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1+l-r,1+r-l\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right)
\]
when $r>l$.

@@

The Kochen-Specker theorem asserts that there is no function~$v$
assigning a definitely value to all observables such that \begin{subequations}
\label{eq:Kochen-Specker-Rules} 
\begin{eqnarray}
 &  & v\left(\mathbf{O}_{1}\right)+v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}+\mathbf{O}_{2}\right)\textrm{ and}\\
 &  & v\left(\mathbf{O}_{1}\right)v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}\mathbf{O}_{2}\right)
\end{eqnarray}
\end{subequations} if $\mathbf{O}_{1}$ and $\mathbf{O}_{2}$ are
commuting observables. Suppose that there is a function~$v$ satisfying
these two properties, it must send an observable to its eigenvalues
so that the following observables~\citep{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{array}{ccc}
\mathbb{1}\otimes\sigma_{z} & \sigma_{z}\otimes\mathbb{1} & \sigma_{z}\otimes\sigma_{z}\\
\sigma_{x}\otimes\mathbb{1} & \mathbb{1}\otimes\sigma_{x} & \sigma_{x}\otimes\sigma_{x}\\
\sigma_{x}\otimes\sigma_{z} & \sigma_{z}\otimes\sigma_{x} & \sigma_{y}\otimes\sigma_{y}
\end{array}\label{eq:MerminSquare}
\end{equation}
must be sent to $1$ or $-1$. In each row and column, operators commutes,
and each operator is the product of the two others, except in the
third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a function~$v$ such that
\begin{equation}
\begin{array}{ccc}
v\left(\mathbb{1}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\mathbb{1}\right) & v\left(\sigma_{z}\otimes\sigma_{z}\right)\\
v\left(\sigma_{x}\otimes\mathbb{1}\right) & v\left(\mathbb{1}\otimes\sigma_{x}\right) & v\left(\sigma_{x}\otimes\sigma_{x}\right)\\
v\left(\sigma_{x}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\sigma_{x}\right) & v\left(\sigma_{y}\otimes\sigma_{y}\right)
\end{array}\label{eq:MerminSquare-values}
\end{equation}
satisfying equation~(\ref{eq:Kochen-Specker-Rules}).

>>>>

In conventional quantum theory, when we measure a property $\mathbf{O}$
of a physical system whose state is represented by a density matrix~$\rho$,
the probability to obtain the value $\lambda$ of that property is determined by the
Born rule $\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{\lambda}\right)=\Tr\left(\rho P_{\lambda}\right)$,
where $P_{\lambda}$ denotes the projector onto the eigenspace corresponding
to the eigenvalue $\lambda$ of the Hermitian operator
$\mathbf{O}$~\citep{Born1983,peres1995quantum,544199,Jaeger2007}. 
Alternatively, one may say that the state $\rho$ determines the 
probability measure $\mu_{\rho}^{\mathrm{\mathrm{B}}}$ defined as a
function from the set $\events$ of all possible projectors $P$ onto the interval 
$[0,1]$, and satisfies the following probability axioms~
\citep{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}: 
\begin{itemize}
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\mathbb{0})=0$. 
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}(\mathbb{1})=1$. 
\item $\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\mathbb{1}-P\right)=
1-\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P\right)$. 
\item For a set of mutually orthogonal projectors $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_iP_j=\delta_{ij}P_i$,
we have 
\begin{equation}
\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\sum_{i=1}^{N}P_{i}\right)=
\sum_{i=1}^{N}\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{i}\right)
\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
satisfying the above axioms, Gleason's theorem asserts that there
exists a unique state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{\mathrm{B}}}$
in a Hilbert space of dimension $d\geq3$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.

\paragraph{Quantum Interval-valued Probability Measures ({\rm IVPMs})}

To formalize the idea of finite precision measurement, we next introduce 
the notion of quantum IVPM~$\bar{\mu}$. The latter replaces 
the infinitely precise interval $[0,1]$ by a finite collection $\mathscr{I}$ of
intervals. Since the intervals $\imposs=\left[0,0\right]$ and $\necess=\left[1,1\right]$ 
should be included in $\mathscr{I}$, and arithmetic operations on intervals satisfy
\begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & \left[1,1\right]-\left[l,r\right]=\left[1-r,1-l\right], \textrm{ and}\\
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}] ,
\end{eqnarray}
\end{subequations}
$\bar{\mu}:\events\rightarrow\mathscr{I}$ is defined by the axioms:
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item 
$
\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Complement}
$
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
$P_iP_j=\delta_{ij}P_i$, 
we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
Notice that Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
relaxes the equality in Eq.~(\ref{eq:QuantumProbability-Addition})
to inclusion because the equality is usually too strong to hold. For
example, if we consider $\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $
and $\bar{\mu}\left(P\right)=\left[l,r\right]$, by Eq.~(\ref{eq:QuantumInterval-valuedProbability-Complement})
and (\ref{eq:interval-operations}), we have 
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1+l-r,1+r-l\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right)
\]
when $r>l$.

@@

The Kochen-Specker theorem asserts that there is no function~$v$
assigning a definite value to all observables such that \begin{subequations}
\label{eq:Kochen-Specker-Rules} 
\begin{eqnarray}
 &  & v\left(\mathbf{O}_{1}\right)+v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}+\mathbf{O}_{2}\right)\textrm{ and}\\
 &  & v\left(\mathbf{O}_{1}\right)v\left(\mathbf{O}_{2}\right)=v\left(\mathbf{O}_{1}\mathbf{O}_{2}\right)
\end{eqnarray}
\end{subequations} if $\mathbf{O}_{1}$ and $\mathbf{O}_{2}$ are
commuting observables. Suppose that there is a function~$v$ satisfying
these two properties, it must send an observable to its eigenvalues
so that the following observables~\citep{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{array}{ccc}
\mathbb{1}\otimes\sigma_{z} & \sigma_{z}\otimes\mathbb{1} & \sigma_{z}\otimes\sigma_{z}\\
\sigma_{x}\otimes\mathbb{1} & \mathbb{1}\otimes\sigma_{x} & \sigma_{x}\otimes\sigma_{x}\\
\sigma_{x}\otimes\sigma_{z} & \sigma_{z}\otimes\sigma_{x} & \sigma_{y}\otimes\sigma_{y}
\end{array}\label{eq:MerminSquare}
\end{equation}
must be sent to $1$ or $-1$. In each row and column, operators commutes,
and each operator is the product of the two others, except in the
third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a function~$v$ such that
\begin{equation}
\begin{array}{ccc}
v\left(\mathbb{1}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\mathbb{1}\right) & v\left(\sigma_{z}\otimes\sigma_{z}\right)\\
v\left(\sigma_{x}\otimes\mathbb{1}\right) & v\left(\mathbb{1}\otimes\sigma_{x}\right) & v\left(\sigma_{x}\otimes\sigma_{x}\right)\\
v\left(\sigma_{x}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\sigma_{x}\right) & v\left(\sigma_{y}\otimes\sigma_{y}\right)
\end{array}\label{eq:MerminSquare-values}
\end{equation}
satisfying equation~(\ref{eq:Kochen-Specker-Rules}).
